---
title: "How familiarity warps representation in face space — Analysis"
author: "Vassiki Chauhan, Maria Ida Gobbini"
output: pdf_document
---
**Experimental paradigm:**

Fixation cross for either 500 or 700 ms
Morphed face (10% to 90% in steps of 10%) for 1000 ms
Original faces for 2AFC displayed until response

**Counterbalancing parameters:**

6 blocks total. 3 blocks from familiar faces,
3 blocks from unfamiliar faces

108 trials per block, each identity at each morph
percentage presented 4 times in a block, 12 times
over the course of the experiment

Consecutive trials from different identities, block
order counterbalanced across subjects



Loading Libraries...
```{r, echo=FALSE}
require('ggplot2')
require('lme4')
require('car')
require('dplyr')
library('lmerTest')
```


# Reaction Times


```{r}

data <- read.csv('~/Desktop/CatPercep/Data/data.csv')

data$cond <- as.character(data$cond)

# removing outliers from reaction times
data <- filter(data, rt > 0.15 & rt < 5)

# indexing separate stimulus conditions
for (x in 1:length(data$id)){
  data$stim[x] <- paste(data$cond[x],'_',data$block[x],sep="")
}

# defining labels to flip unfamiliar-unfamiliar blocks
stim_conditions <- c("f2_unfam","m2_unfam","m1_unfam")
flip_conditions <- paste(stim_conditions, "_flip", sep="")
conditions <- data.frame(c(stim_conditions[1],stim_conditions[2],stim_conditions[3]),
                         c(stim_conditions[1],flip_conditions[2],stim_conditions[3]),
                         c(stim_conditions[1],flip_conditions[2],flip_conditions[3]),
                         c(flip_conditions[1],stim_conditions[2],flip_conditions[3]),
                         c(flip_conditions[1],flip_conditions[2],stim_conditions[3]),
                         c(flip_conditions[1],flip_conditions[2],flip_conditions[3]),
                         c(stim_conditions[1],stim_conditions[2],flip_conditions[3]),
                         c(flip_conditions[1],stim_conditions[2],stim_conditions[3]),
                         fix.empty.names = FALSE   ) 
conditions <- as.data.frame(t(conditions))
colnames(conditions) <- c("stim1","stim2","stim3")
rep_labels <- c("c1","c2","c3","c4","c5","c6","c7","c8")
conditions$labels <- rep_labels

# computing all possible combinations of flipped unfamiliar blocks
for (i in 1:length(stim_conditions)){
  flip_label <- flip_conditions[i]
  data_to_flip <- filter(data,stim == stim_conditions[i])
  for (j in 1:length(data_to_flip$id)){
    if (data_to_flip$ans[j] == 'perc100'){
      #data_to_flip$ans[j] = 'perc0'
      data_to_flip$ans[j] = 'perc0'
    }else{
      data_to_flip$ans[j] = 'perc100'
    }
    data_to_flip$morph[j] = 100-data_to_flip$morph[j]
    data_to_flip$famfaceloc[j] <- ifelse(data_to_flip$famfaceloc[j] == 'left','right','left')
    data_to_flip$stim[j] <- flip_label
  }
  # adding flipped trials to main data frame
  data <- bind_rows(data,data_to_flip)
}

data$acc <- ifelse((data$morph < 50 & data$famfaceloc != data$resp) | 
                     (data$morph > 50 & data$famfaceloc == data$resp),1, 0)

# subsetting correct trials
data_correct <- filter(data,acc == 1)
```


## Testing different models for reaction times


```{r}
# we need to scale the variable morph and treat it as a continuous variable
data_correct$morph <- as.factor(data_correct$morph)
data_correct$morph_sc <- as.numeric(as.character(data_correct$morph))
morph_scaled <- scale(data_correct$morph_sc)
data_correct$morph_scale <- as.numeric(morph_scaled)

contrasts(data$block) <- contr.sum

model_1_rt <- lmer(log(rt) ~ block*morph_scale +
               (1 + block | id) +
               (1 + block | stim),
             data = data_correct, REML = F)

model_2_rt <- lmer(log(rt)~block*morph_scale + (1+block|id) + 
                   (1|stim), data = data_correct, REML = F)

anova(model_1_rt,model_2_rt)
```


Random intercept of block with stimulus condition does not improve model fit, 
and our first model fails to converge.


```{r}
model_3_rt <- lmer(log(rt)~block*morph_scale + (1|id) + 
                   (1|stim), data = data_correct, REML = F)

anova(model_2_rt,model_3_rt)

```
  
  
 However, the random effect of stimulus condition clearly improves our model. This will be the model
 we report statistics on! Refitting the model with restricted maximum likelihood and looking at 
 which of the fixed effects are significant. 


## Model Output


```{r}
model_rt <- update(model_2_rt, REML=T)
Anova(model_rt,type = 3)
```


# Percentage "Identity B" Responses 


```{r}
# please clean your workspace and re enter the data!!
rm(list=ls())

data <- read.csv('~/Desktop/CatPercep/Data/data.csv')

data$cond <- as.character(data$cond)

# removing outliers again 
data <- filter(data, rt > 0.15 & rt < 5)

for (x in 1:length(data$id)){
  data$stim[x] <- paste(data$cond[x],'_',data$block[x],sep="")
}

stim_conditions <- c("f2_unfam","m2_unfam","m1_unfam")
flip_conditions <- paste(stim_conditions, "_flip", sep="")
conditions <- data.frame(c(stim_conditions[1],stim_conditions[2],stim_conditions[3]),
                         c(stim_conditions[1],flip_conditions[2],stim_conditions[3]),
                         c(stim_conditions[1],flip_conditions[2],flip_conditions[3]),
                         c(flip_conditions[1],stim_conditions[2],flip_conditions[3]),
                         c(flip_conditions[1],flip_conditions[2],stim_conditions[3]),
                         c(flip_conditions[1],flip_conditions[2],flip_conditions[3]),
                         c(stim_conditions[1],stim_conditions[2],flip_conditions[3]),
                         c(flip_conditions[1],stim_conditions[2],stim_conditions[3]),
                         fix.empty.names = FALSE   ) 
conditions <- as.data.frame(t(conditions))
colnames(conditions) <- c("stim1","stim2","stim3")
rep_labels <- c("c1","c2","c3","c4","c5","c6","c7","c8")
conditions$labels <- rep_labels

# flipping labels
for (i in 1:length(stim_conditions)){
  flip_label <- flip_conditions[i]
  data_to_flip <- filter(data,stim == stim_conditions[i])
  for (j in 1:length(data_to_flip$id)){
    if (data_to_flip$ans[j] == 'perc100'){
      #data_to_flip$ans[j] = 'perc0'
      data_to_flip$ans[j] = 'perc0'
    }else{
      data_to_flip$ans[j] = 'perc100'
    }
    data_to_flip$morph[j] = 100-data_to_flip$morph[j]
    data_to_flip$stim[j] <- flip_label
  }
  data <- bind_rows(data,data_to_flip)
}

```


## Testing different models for percentage "Identity B responses"


```{r}
data$morph <- as.factor(data$morph)
data$block <- as.factor(data$block)

contrasts(data$block) <- contr.sum

data$morph_sc <- as.numeric(as.character(data$morph))
morph_scaled <- scale(data$morph_sc)
data$morph_scale <- as.numeric(morph_scaled)

md1 <- glmer(ans ~ block*morph_scale +
               (1 + block | id) +
               (1 + block | stim),
             family=binomial,
             data = data,
             control=glmerControl(optimizer="bobyqa")) 

md2 <- glmer(ans ~ block*morph_scale +
               (1 + block | id) +
               (1 | stim),
             family=binomial,
             data = data,
             control=glmerControl(optimizer="bobyqa")) 

anova(md1,md2)
```
  
  
Again, the random intercept due to block does not improve model fit. 


```{r}
md3 <- glmer(ans ~ block*morph_scale +
               (1 + block | id),
             family=binomial,
             data = data,
             control=glmerControl(optimizer="bobyqa")) 

anova(md2,md3)
```


Clearly, `md2` will be our final statistical model

## Model Output
```{r}
Anova(md2, type=3)
summary(md2)
```